---
title: "introduction_to_MutSignature"
author: "Duccio Pagliai"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{introduction_to_MutSignature}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
# Install package
#devtools::install_github("ducciopagliai/MutSignature")
library(MutSignature)
```

## Description
`MutSignature` is a tool designed to analyze genomic loci and assess their mutational signatures based on COSMIC mutational patterns. 
The use of the package is demonstrated with an example.


## Step 1: Extract DNA Sequences from Genomic Loci
The extract_sequences function retrieves DNA sequences from a reference genome given a set of loci in the format chr:start-end.

```{r}
library(BSgenome.Hsapiens.UCSC.hg38)
library(decompTumor2Sig)
# Define loci
loci <- c("chr1:1000000-1000100", "chr2:2000000-2000100")

# Extract sequences
sequence_df <- extract_sequences(loci, BSgenome.Hsapiens.UCSC.hg38)
print(sequence_df)
```

## Step 2: Count Triplets
Use the count_triplets_per_locus function to count triplets in each sequence or count_triplets_aggregated to count triplets across all loci.

```{r}
# Count triplets for each locus
triplet_counts <- count_triplets_per_locus(sequence_df)
print(triplet_counts)

# Aggregate triplet counts across all loci
global_triplet_counts <- count_triplets_aggregated(triplet_counts)
print(global_triplet_counts)
```


## Step 3: Map Triplets to COSMIC format
Use the map_to_cosmic function to map the triplet counts to the COSMIC mutational signatures.

```{r}
# Define the path to the COSMIC file
cosmic_path <- system.file("extdata", "COSMIC_v3.4_SBS_GRCh38 2.txt", package = "MutSignature")

# Map counts per locus to COSMIC format
cosmic_counts_per_locus <- map_to_cosmic(triplet_counts, cosmic_path)
print(cosmic_counts_per_locus)

# Map aggregated counts to COSMIC format
cosmic_counts_aggregated <- map_to_cosmic(global_triplet_counts, cosmic_path)
print(cosmic_counts_aggregated)
```


## Step 4: Decompose Mutational Signatures
Use the decompose_signatures function to determine the contribution of each COSMIC signature to the observed mutations.

```{r}
# Decompose per locus
result_per_locus <- decompose_signatures(cosmic_counts_per_locus, cosmic_path)
print(result_per_locus)

# Decompose aggregated counts
result_aggregated <- decompose_signatures(cosmic_counts_aggregated, cosmic_path)
print(result_aggregated)
```


## Additional resources
- COSMIC Signatures:
https://cancer.sanger.ac.uk/signatures/








